---
title: "Makarova_labaR"
author: "Makarova Mariia"
date: "04 11 2020"
output: html_document
runtime: shiny
params:
  n: 32038
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Load data

Loading data from file named zipIncome.txt

  * sep - the field separator character
  * header - a logical value indicating whether the file contains the names of the variables as its first line.
  * nrows - integer: the maximum number of rows to read in.
  
  x <- 1

  x <- c(1, 2)
  
rbind
  
 1 2 3

 4 5 6 

 7 8 9
   
cbind   
   
 1 4 7 

 2 5 8

 3 6 9
   
```{r loading data}
table <- read.table('/cloud/project/zipIncome.txt', sep='|', header=TRUE)
```

  A variable "table" is a data frame.

```{r checking class}
class(table)
```

```{r, echo=FALSE, fig.align='center'}
numericInput("n", "How many first rows to show?", 5)
renderTable({
  head(table, input$n)
})
```

## Renaming columns

Changing the column names of our data frame so that zip_Prefix becomes zipCode and meanhouseholdincome becomes income.

```{r renaming columns}
colnames(table)[1] <- "zipCode"
colnames(table)[2] <- "income"
```

## Mean and Median Average Incomes

Analyzing the summary of our data. Here we can see the mean and median average incomes.
Summary() provides the minimum, maximum, median, mean and the 1st and 3rd quartile values.

```{r showing summary}
summary(table)
mean(table$income, na.rm=TRUE)
median(table$income, na.rm=TRUE)
```
## Plotting a scatter plot of the data

```{r plotting a scatter plot of the data}
plot(table)
```

## Filtering the data

In order to omit outliers, we create a subset of the data so that: $7,000 < income < $200,000

```{r filtering the data}
filtered_data <- table[table$income >= 7000 & table$income <= 200000,]
summary(filtered_data)
mean(filtered_data$income, na.rm=TRUE)
```
## Simple Boxplot (Box and Whisker Plot)

In the box plot you created, notice that all of the income data is pushed towards the bottom of the graph because most average incomes tend to be low.
* str(x) - internal structure of object x .
* We can generate factor levels by using the gl() function.

```{r creating simple boxplot}
str(filtered_data)

boxplot(income~as.factor(zipCode), 
        data=filtered_data,
        main = "Average Household Income by Zip Code",
        xlab = "Zip Codes",
        ylab = "Income"
)
```

Create a new box plot where the y-axis uses a log scale.  Be sure to add a title and label the axes
log parameter in boxplot function for y-axis 

```{r creatng a new simple boxplot}
boxplot(income~as.factor(zipCode), 
        data=filtered_data,
        main = "Average Household Income by Zip Code",
        xlab = "Zip Codes",
        ylab = "Income",
        log= 'y'
)
```

 Make a ggplot that consists of just a scatter plot (диаграмма рассеяния) using the function geom_point() with position = “jitter”  so that the data points are grouped by zip code.  Be sure to use ggplot’s function for taking the log10 of the y-axis data.  (Hint: for geom_point, have alpha=0.2).
 
 * data - ggplot builds a plot by using this data
 * aes - begins a plot that you finish by adding layers to. Add one geom function per layer
 * geom_point - geom points all over the plot
 
 Position adjustments - position adjustments determine how to arrange geoms that would otherwise occupy the same space.
 
 * e + geom_point(position="jitter) - add random noise to X and Y position of each element to avoid overplotting
 
 * scale_y_log10() - plot y on log10
 
```{r }
ggplot(data=filtered_data, aes(x=as.factor(zipCode), y=income)) + 
  geom_point(aes(x = as.factor(zipCode)), position = "jitter", alpha = 0.2) +
  scale_y_log10()
```

Create a new ggplot by adding a box plot layer to your previous graph.  To do this, add the ggplot function geom_boxplot().  Also, add color to the scatter plot so that data points between different zip codes are different colors.  Be sure to label the axes and add a title to the graph.

```{r }
ggplot(data=filtered_data, 
       aes(x=as.factor(zipCode), y=income)) + 
        geom_point(aes(color = as.factor(zipCode)), position = "jitter", alpha = 0.2) +
        scale_y_log10() +
        geom_boxplot(alpha = 0.1, outlier.size = -Inf) +
        labs(color="Region", x="Zip Code", y="Income", title="Average Income by Zip Code")
```